Title of Paper

### 2017-02-16 paper.tex
update data ordering discussion

\subsection{Data Ordering and Vectorization Patterns}
For a chemical kinetic model with species $k$ for the $j$-th thermo-chemical state to be evaluated, $[C_k]_j$ there are two choices for data-storage in memory.
In the ``C'' (C row-major) format, species concentrations for $j$-th thermo-chemical state are stored sequentially in memory.
Alternatively, in the ``F'' (Fortan column-major) format species concentrations a single species for over all thermo-chemical states are adjacent in memory.
In Sec.~\ref{S:results} the effect of this ordering will be shown to have a large performance impact on the program runtime.

In this work, a \textit{shallow}-SIMD vectorization has been implemented, entailing each SIMD-lane evaluating chemical source terms for an thermo-chemical state independently.
In future efforts, a so called \textit{deep}-SIMD vectorization will be pursued allowing vector-units to cooperate to evaluate species rates for a single thermo-chemical state.
A deep-vectorization may result in SIMD-\textit{waste} in this context---similar to thread-divergence in SIMT-processing---caused by different SIMD-lanes executing different instructions (e.g. from differing for-loop bounds).
However a deep-vectorization will likely have higher cache hit-rates than a shallow-vectorization using the ``C'' memory layout as successive SIMD-lanes would access directly memory adjacent locations.
Additionally, past studies demonstrate the merits of deep-SIMT vectorization on GPUs~\cite{Sewerin20151375}.

>>>>

\subsection{Data Ordering and Vectorization Patterns}
\label{S:data}
When storing arrays for a chemical kinetic model, e.g. the concentration of species $k$ for the $j$-th of thermo-chemical state $[C_k]_j$ there are two data-ordering choices.
In the ``F'' (Fortran column-major) format, the concentrations of a single species $k$ over all $N$ thermo-chemical states are adjacent: $[C_k]_j, [C_k]_{j + 1}\ldots[C_k]_{N}$;
conversely in the ``C'' (C row-major) format, the species concentrations for the $j$-th thermo-chemical state: $[C_k]_j, [C_{k+1}]_j\ldots$ are stored sequentially in memory;
This ordering---along with the device (CPU, GPU, etc.) in question---have a large effect on the performance of SIMD\slash SIMT-vectorized algorithms.

In this work a \textit{shallow}-SIMD vectorization, in which each SIMD-lane evaluates chemical source terms for a different thermo-chemical state, has been implemented.
If the data is stored in ``F''-order, the SIMD-lanes will access sequential locations in memory however the next species' concentration---for example---will be $N$ memory locations away, potentially resulting in more cache-misses on the CPU.
In future efforts, a so called \textit{deep}-SIMD vectorization will be pursued, in which a vector-unit cooperates to evaluate chemical source terms for a single thermo-chemical state.
In this context, a deep-vectorization may result in SIMD-\textit{waste}---similar to thread-divergence in GPU-processing---caused by different SIMD-lanes executing different instructions (e.g. from differing for-loop bounds).
However, SIMD-lanes in a ``C''-ordered deep-vectorization access sequential memory locations and further the data locality of e.g., species concentrations, will likely result in improved CPU cache-hit rates.
Past studies additionally demonstrate promising results for deep-SIMT vectorization on GPUs~\cite{Sewerin20151375}.
In Sec.~\ref{S:results} the effect of this ordering on the shallow-SIMD vectorization performance will be studied.
